---
title: "mcroarray data analysis"
author: "Jiang Nan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 1. Get packages
we need to use the limma model to srceen the DEG

```{r}

library(limma) # load limma model 
library(annotate) # annote microarray
library(impute)  # for deal with missing value
library(ggplot2)  #加载ggplot2包
library(pheatmap) #加载pheatmap包
```

### parameters
```{r}
main_path ='C:/Users/DELL/Desktop/WGCNA-cmap-cancer/data/GSE14520/'
eset_file_path="GSE14520-GPL3921_series_matrix.txt"
group_file_path="group-of-GPL3921_series.csv"
anotation_Plaform='hthgu133a'  #check GEO see the bioc_backage name of the microarray
adj_PVal=0.05
logFC=1

```

### Load data and sample comparation vector

```{r}
setwd(main_path)
eset<-read.table(eset_file_path, sep='\t', header=T, comment.char = "!", blank.lines.skip = TRUE, row.names=1) #load gene expression matrix, omit those comment and blank rows

#head(eset_txt)  # see if the data have already log2
#eset_txt<-log2(eset_txt) # if the data has already been log2 then, skip this line.
```

### check the data deviation

```{r}
boxplot(eset, col="blue", xaxt='n', outline=F)

```
### check the data deviation is too big the normalize the matrix

```{r}
eset2 = normalizeBetweenArrays(as.matrix(eset),method="scale")

boxplot(eset2,col = "red",xaxt = "n",outline = F)
```

### Probe ID into Gene symbol 中位数合并表达值（对平台文件进行整理）

```{r}
# probe_exp<-read.table(eset_file_path, sep='\t', header=T, comment.char = "!", blank.lines.skip = TRUE, row.names=1)
# get annotation info
affydb<-annPkgName(anotation_Plaform,type="db")
#BiocManager::install(affydb)
library(affydb,character.only=TRUE)
eset$symbols<-getSYMBOL(rownames(eset),affydb)

# merge the mutiple value with same gene symble
index<-which(!is.na(eset$symbols))
eset_matrix<-eset[index,]
#write.csv(exp_matrix,file="exp_matrix.csv",quote=F)
geneidfactor<-factor(eset_matrix$symbols)
#head(geneidfactor)
eset_matrix<- eset_matrix[,-length(eset_matrix)]
#head(exp_matrix)
gene_eset_matrix<-apply(eset_matrix,2,function(x) tapply(x,geneidfactor, median))

write.csv(gene_eset_matrix,file=paste(main_path,"gene_eset_matrix.csv",sep = ""),quote=F)
```

### DEAL WITH missing values 补充缺失值 
#最近邻居法（KNN，k-Nearest Neighbor）法：此方法是寻找和有缺失值的基因的表达谱相似的其他基因，通过这些基因的表达值（依照表达谱相似性加权）来填充缺失值

```{r}
#读取表达值
#gene_eset_matrix<-read.table("gene_eset_matrix.csv",header=T,sep="\t",row.names=1)
gene_eset_matrix<-as.matrix(gene_eset_matrix)
#KNN法计算缺失值
imputed_gene_exp<-impute.knn(gene_eset_matrix,k=10,rowmax=0.5,colmax=0.8,maxp=3000,rng.seed=362436069)
#读出经过缺失值处理的数据
GeneExp<-imputed_gene_exp$data
#写入表格
genesymbol<-rownames(GeneExp)
gene_eset_matrix<-cbind(genesymbol,GeneExp)
write.csv(GeneExp,file=paste(main_path,"gene_eset_matrix2.csv",sep = ""),quote=F)
```


### Set sample comparation vector 分组

make sure in the group file the cancer value is 1, the non-cancer value is 0, and column name is "type"

```{r}
setwd(main_path)

groups<-read.csv(group_file_path, sep=',', header=T) #load gene expression

groups<-factor(groups$type)
#head(groups)

```

### Use the limma model get gene expression table

```{r}
setwd(main_path)
design<-model.matrix(~-1+groups)
contrast.matrix<-makeContrasts(contrasts="groups1-groups0",levels=design) #构建对比模型，并存储于变量contrast.matrix

fit<-lmFit(gene_eset_matrix,design) #The linear model fit 
fit1<-contrasts.fit(fit,contrast.matrix) # contrast computing
fit2<-eBayes(fit1)    # Bayes test
allgene<-topTable(fit2,coef="groups1-groups0",n=nrow(fit2),sort.by="p") #Generate a checklist of all the genes

write.csv(allgene,file="allgene DE result.csv",quote=F)
```

### Classify DEGs


```{r}
setwd(main_path)
allgene$type<-NA
#给allgene增加一列type，并给数据赋值为NA
for(i in 1:nrow(allgene))if(allgene[i,1]>logFC & allgene[i,5]<adj_PVal){allgene[i,7]<-"up"}else if(-allgene[i,1]>logFC & allgene[i,5]<adj_PVal) {allgene[i,7]<-"down"}else{allgene[i,7]<-"normal"}
#通过循环语句，将基因类型分为三类（logFC>1.5，adj.p.Val<0.01的为上调基因up，logFC<-1.5,adj.p.Val<0.01的为下调基因down，其他为normal）
#write.csv(allgene,file="allgene2.csv",quote=F)
#将新的所有基因的检测报告，输出为csv格式。
gene_up<-allgene[allgene$type=="up",]
#筛选上调差异基因，并存储于变量gene_up中
gene_down<-allgene[allgene$type=="down",]
#筛选下调差异基因，并存储于变量gene_down中
gene_dif<-allgene[allgene$type!="normal",]
#筛选非差异基因，并存储于变量gene_dif中
write.csv(gene_up,file="DE_gene_up.csv",quote=F)
#将筛选的上调基因检测报告，保存为csv格式
write.csv(gene_down,file="DE_gene_down.csv",quote=F)
#将筛选的下调基因检测报告，保存为csv格式
write.csv(gene_dif,file="DE_gene_dif.csv",quote=F)
#将筛选的差异基因检测报告，保存为csv格式

```

### volcano map and heat map

```{r}
setwd(main_path)
threshold<-factor(allgene$type)
#将type转化为因子向量，并存储于变量threshold中，用来区分火山图差异基因颜色鑹?
pdf(file="volcano.pdf",width=10,height=10)

ggplot(allgene,aes(x=logFC,y=-log10(adj.P.Val),colour=threshold))+xlab("log2FC")+ylab("-log10adj.P-Value")+ggtitle("Volcano Plot")+theme(plot.title=element_text(hjust=0.5))+geom_point()+geom_vline(xintercept = c(-logFC,logFC))+geom_hline(yintercept = -log10(adj_PVal))+theme(panel.grid =element_blank())+scale_colour_manual(values=c("green", "black", "red"))+expand_limits(x=c(-8,8))
dev.off() # close the device 保存关闭绘图设备
```

### heat map

```{r}
setwd(main_path)
gene_dif_exp<-gene_eset_matrix[rownames(gene_dif),]
#提取差异基因的基因表达值
write.csv(gene_dif_exp,file="gene_dif_exp.csv",quote=F)
#将筛选的差异基因表达，保存为csv格式
tiff(file="gene_dif_pheatmap.tif",res=300,units='in',width=30,height=30)

pheatmap(gene_dif_exp,color=colorRampPalette(c("green","black","red"))(100),fontsize_row=3,scale="row",border_color=NA)
dev.off() #保存，关闭绘图设备

```




```{r}

```
